Legacy Data Confounds Modern Genomics Studies

### 2019-08-08 LatexManuscript/MBE/legacy.tex
chromosomal enrichment

Thus, this mutational signal is heavily enriched in $Q$-associated SNPs, but residual signal remains in non-significant SNPs, presumably because many rare alleles found in individuals with low $Q$ remain unidentifiable using association techniques due to lack of power (Supplementary Figure \ref{histogram_of_sig_snps}).
The removal of individuals with $Q$ below 30 successfully removes the *AC${\rightarrow}$*CC signal, however other signals identified by Harris and Pritchard appear unchanged (Figure \ref{MutSpect}).
For population genetic analyses sensitive to the accumulation of rare variants, the removal of individuals with low $Q$ appears preferable to filtering specific low-quality SNPs. 
For other analyses where quality of imputation matters, identifying $Q$-associated variants may be preferable.

@@

$Q$-associated variants are distributed across the genome, with chromosome 1 showing an excess of such variants, and other chromosomes being relatively uniform (Figure \ref{CountPerChrom}).
At a 10kb scale, we also see rather uniform distribution with a small number of regions showing an enrichment for such variants (Figure \ref{VariantDensity}).
An outlying 10kb region in chromosome 17 (bases : 22,020,000 to  22,030,000) has 35 $Q$-associated variants.
Distribution of association statistics in this region is provided in Figure \ref{Top10}.
By contrast, variants that do not pass the 1kGP strict mask are more unevenly distributed across the genome (Figure \ref{VariantDensity}).

@@

@@

>>>>

Thus, this mutational signal is heavily enriched in $Q$-associated SNPs, but residual signal remains in non-significant SNPs, presumably because many rare alleles found in individuals with low $Q$ remain unidentifiable using association techniques due to lack of power (Supplementary Figure \ref{histogram_of_sig_snps}).
\todo{The removal of individuals with $Q$ below 30 successfully removes the *AC${\rightarrow}$*CC signal enrichment in the Japanese, however other signals identified by Harris and Pritchard appear unchanged including the continental enrichment of *AC${\rightarrow}$*CC signal in East Asians compared to Africans as reported by Harris and Pritchard and replicated in the SGDP (Figure \ref{MutSpect} and \ref{MutSpect_continental}).}
For population genetic analyses sensitive to the accumulation of rare variants, the removal of individuals with low $Q$ appears preferable to filtering specific low-quality SNPs. 
For other analyses where quality of imputation matters, identifying $Q$-associated variants may be preferable.

@@

$Q$-associated variants are distributed across the genome, with chromosome 1 showing an excess of such variants, and other chromosomes being relatively uniform (Figure \ref{CountPerChrom}).
\todo{Chromosome 1 shows enrichments in the  *AC${\rightarrow}$*CC signal compared to other chromosomes despite accounting for chromosomal length (Figure \ref{MutEnrichment_chrom}).} 
At a 10kb scale, we also see rather uniform distribution with a small number of regions showing an enrichment for such variants (Figure \ref{VariantDensity}).
An outlying 10kb region in chromosome 17 (bases : 22,020,000 to  22,030,000) has 35 $Q$-associated variants.
Distribution of association statistics in this region is provided in Figure \ref{Top10}.
By contrast, variants that do not pass the 1kGP strict mask are more unevenly distributed across the genome (Figure \ref{VariantDensity}).

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=10cm,keepaspectratio]{mutational_enrichment_per_chrom.jpg}
\caption{Mutation spectrum enrichment of  $Q$-associated variants across each chromosome.
Chromosome 1 has a strong enrichment in *AC${\rightarrow}$*CC mutational pattern.}  
\label{MutEnrichment_chrom}
\end{figure}

@@

\begin{figure}[tbp]
\includegraphics[width=11cm,keepaspectratio]{CutOffID_QUAL30_heatmap_v_AFR_nosingle.jpg}
\caption{Comparing continental mutational signatures after removing $Q$-associated variants and after removing individuals with low $Q$.  
The  *AC${\rightarrow}$*CC mutational signature in East Asian (EAS) Population remains despite removing individuals with average quality per mapped bases $Q$ below a threshold of 30.
Points indicate significant mutation spectrum contrasts with $p <10^{-5}$}
\label{MutSpect_continental}
\end{figure}

### 2019-08-08 PlottingScripts/ProcessPlotData.Rmd
chromosomal enrichment

return(binCount)
}

>>>>

return(binCount)
}
```

##per chromosome
```{r}
ChromLength <- fread(paste0(DataPath,
                            'GRCh37_ChromLength.txt'),
                     sep='\t',
                     col.names = c('Chr','Length'))

snp_list <- list()
snp_combined <- data.frame()
maxFreq <- 0
for(chrom in 1:22) {
  snp_list[[chrom]] <- 
    processContext(SNP %>% filter(Chr == chrom))
  if(max(snp_list[[chrom]]$Freq) > maxFreq){
    maxFreq <- max(snp_list[[chrom]]$Freq)
  }
  snp_list[[chrom]]$Chr <- chrom
  snp_combined <- rbind(snp_combined,snp_list[[chrom]])
}

plt_list <-list()
for(chrom in 1:22) {
  plt_list[[chrom]] <- 
    plotContext(snp_list[[chrom]],paste('Chrom',chrom)) + 
                theme(legend.position = "none")
}

plot_grid(plt_list[[1]] , plt_list[[2]], plt_list[[3]], plt_list[[4]], plt_list[[5]], plt_list[[6]], plt_list[[7]], plt_list[[8]], plt_list[[9]], plt_list[[10]], plt_list[[11]], plt_list[[12]], plt_list[[13]], plt_list[[14]], plt_list[[15]], plt_list[[16]], plt_list[[17]], plt_list[[18]], plt_list[[19]], plt_list[[20]],plt_list[[21]], plt_list[[22]], ncol=2)

ggsave(paste0(figurePath,
              'mutational_enrichment_per_chrom.jpg'),
              height=24, 
              width=12)

snp_combined$title <- paste(snp_combined$Start,"*",snp_combined$End)

snp_combined <- merge(snp_combined, ChromLength, by = 'Chr')

snp_combined$norm_freq <- snp_combined$Freq / snp_combined$Length

ggplot(snp_combined, 
       aes(x=Mut, 
           y = norm_freq, 
           color = as.factor(Chr)))+
  geom_point(shape =1) +
  geom_text_repel(data = snp_combined %>% 
              filter(Freq > 200),
              show.legend = FALSE,
    aes(label = title,
        hjust = -1))+
  guides(color=guide_legend(title="Chromosome"))

ggplot(snp_combined, 
       aes(x=Start, 
           y = Freq, 
           color = as.factor(Chr)))+
  geom_point()+
  facet_grid(End~Mut) +
  guides(color=guide_legend(title="Chromosome"))+
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20), 
        panel.border = element_rect(colour = "black", 
                                    fill = NA, 
                                    size = 1))
ggsave(paste0(figurePath,
              'mutational_spectrum_per_chrom.jpg'),
              height=10, 
              width=12)

  geom_text_repel(data = snp_combined %>% 
                    filter(Freq > 200),
                  aes(x=Mut, 
                  y = Freq,
                  label = Chr))+
  guides(color = F) +
```
